A Technique for Inter-Catalogue Recommendations

### 2014-06-10 paper2/paper.tex
Fill out Implementation section

\section*{Implementation}
Here are some of the details for implementing the LSI method for constructing
inter-catalogue recommendations.

\subsection*{Random Walks}
Topic-specific random walk starting from each product
Parameters needed and the ones chosen

\subsection*{Singular Value Decomposition}
Subtract mean before hand (why this is good here but not usually)
Number of concepts to use

\subsection*{Concept Mapping}
Issue of normalization - normalize each column (L1 and L2), orthogonalize, map
to sphere
Bigrams

\subsection*{Popularity}
Important - definition, popularity breakdown of graphs
Model of popularity and similarity
Removing popularity from random walks
Adding popularity to nearest neighbors / modifying number of predictions
Baseline popularity methods
Number of products mapped to in different methods

\section*{Performance}
\subsection*{Baselines}
Random, tfidf, one model (explain), most popular

@@

>>>>

\section*{Implementation}
Let us dig deeper into the details of the LSI approach described above. There 
are a number of implementation decisions that, in all, dramatically affect the
performance of this method.

\subsection*{Random Walks}
The first step of the LSI approach is to simulate user sessions by randomly
walking the recommender graphs. There are a few ways this could be implemented.
One method is to randomly assign a random walker a starting node and then allow
it to randomly walk the edges, with a probability of termination at each step.
Running this many times is a straight-forward model of user sessions.

To speed up the simulation, we calculate a series of product-specific stationary
distributions. As with any random walk technique, we first interpret the 
recommender graph as a transition matrix. In this walk, for each product we create a
product-specific transition matrix which has added teleportation edges back the
that product. We calculate the stationary distributions of each of these
matrices, one per product. We use these product-specific stationary
distributions as our user sessions. Since there is only one user session per 
product, we have many fewer sessions than we would with a real dataset. However,
each of these stationary distributions can be interpreted as an aggregation of
an infinite number of random walkers starting at each node and travelling until
a random termination time, as in the original model.

This process requires a couple of parameters. The first is the number of steps used to
approximate each product-specific statinary distribution. The second parameter
is the likelihood of teleporting to the special product. For our purposes we use
30 steps and a teleportation probability of 5\%.

\subsection*{Singular Value Decomposition}
Unlike Principal Compenent Analysis, standard LSI does not include any
preprocessing of the data, such as subtracting the mean from each column of the
product-session matrix. A major reason for this is that LSI is typically applied
to natural language processing problems, where the rows are terms, and there is
one term per word in the vocabulary. This otherwise sparse matrix becomes dense
if the mean is subtracted, making many NLP datasets untractable to work with. 
Our datast, on the other hand, has significantly fewer rows and columns and is
already dense. Therefore, there is little added cost to subtracting the
means. We include this preprocessing step and find that we get better results
because of it.

Another design decision is in the size of the truncation of the matrices learned
from SVD. This number dictates the number of latent concepts learned for LSI.
Too few concepts fail to capture enough detail to make accurate representations
of each product. Too many make mapping between concepts challenging. We found a
good balance at 16 concepts.

[Graphs of concepts]

\subsection*{Concept Mapping}
A tricky operation of this procedure is the concept mapping. We attempt to find
the best linear transformation from one concept space to the other. We do this
by representing each concept as a vector of words in the vocabulary, each
weighted with how well it represents the concept. We construct this by weighing
the tfidf of each word in a product's description by how well that word
represents the concept. Note that products can have a positive or negative value
for each concept, thus the word representation of concpets are also positive and
negative.

With these concept representations, we construct an innner-product matrix. The 
value of a cell (i, j) in the matrix is the dot product of concept i in the first 
space with concept j in the second. The columns and rows of the resulting matrix
represent the relative similarity of the descriptive words between two concepts.
But the scale of this matrix can be very large. If the inner-product matrix is
used without any normalization as the linear transformation between the concept
spaces, the scale of the transformed products will be inconsistent with those
originally in that space.

One way to normalize is to normalize each column of the inner-product matrix by
the L1 norm. This preserves the L1 norm of products after the transformation,
but does not preserve the L2 norm. Similarly, we can normalize the columns by
the L2 norm. Neither of these methods preserve the L2 norm, which would be ideal
as we find the nearest neighbors in the other concept space by their L2
distance.

Another approach is to orthogonalize the inner-product matrix. That is, find the
orthogonal matrix nearest to the inner-product matrix. Nearness is defined by
the NORM FUNCTION on the difference of the two matrices. This orthogonal matrix
does perserve the L2 norm of the vectors it transforms, however the
transformation is an approximation of the ideal inner-product matrix and can
distort the relative mapping from one concept to the others.

A third approach to normalizing aims to preserve the angle of the inner-product
transformation. For this approach, we remove the issue of scale by normalizing
all products in the mapped-to space to be unit lenght. Then, after transforming
the products from the other catalogue, we normmalize them as well. Thus the
scales for both sets of products are the same and the angles of the
inner-product transformation are preserved, at the expense of discarding the
norms of all products in their product spaces. This is equivalent to calculating
the cosine similarity between products without any normalization, as cosine
similarity is indeifferent to scale. We find that this method of normalization
performs the best.

Further gains can be seen if we consider bigrams in the description as well as
the unigrams. Here, we treat bigrams as a separate set dimensions to represent
each concept. These are included in the inner product between two concepts,
weighted compared to the unigrams by a parameter $\beta$. $\beta = 0.3$ improves the
performance of the concept mapping.

\subsection*{Popularity}
The most effective addition to the LSI procedure is incorporating the popularity
of products. We do not know a priori how popular an item is as we do not have
explicit user data. However, we can use the recommendation graphs to model the
popularity of each item. The assumption is that a retailer will recommend
popular products more frequently than unpopular ones.

We imagine that recommender graphs are constructed as follows: A retailer first
builds a similarity matrix, using a collaborative filtering method. In this
matrix, all similarity relationships are symmetric. Next, the retailer modifies
the values of the matrix to include popularity. As a simple model, we say that
the popularity-modified similarity score from Product i to Proudct j is the
simple similarity score at (i, j) multiplied by the popularity of Product j. In
this way, more popular items are more frequently recommended.

We see signs of this in the undirected recommender graphs. Although each product
has at most four outgoing recommendations on macys.com, we see that some
products have over one hundred incoming recommendations. With this in mind, we
approximate the popularity of each product as the logarithm of the degree of
that product in its undirected recommender graph. We choose to model it as the
logarithm because a few of the products tend to have degrees in the hundreds.
Without some way of scaling back these popularities, these few very popular
products tend to take over all recommendations.

[Graph of popularity]

Assuming this model of recommender graph construction, which separates similarity
from popularity, we seek to isolate similarity from the recommender graph. Since
popular products were promoted for popularity, we demote them when walking the
graphs to create user sessions. In practice, we divide the likelihood of walking
an edge by the popularity of the product on the other side of that edge. Thus
the concepts learned on that matrix better represent products independent of
popularity.

Proceeding as before, we find the nearest products in concept space. However,
before deeming the nearest neighbors to be the best recommendations, we divide
the distance to each product by its popularity. Since we model distance in 
concept space to be inversely proportional to similarity, this reincorporation
of popularity is analagous to multiplying similarity by popularity, as done in
the original construction of the recommender graph.

A final way in which we incorporate popularity is in the number of
recommendations alotted per product. Since we model the recommender graph as an
undirected graph, different nodes have different degrees. A product that is
popular within its catalogue is likely to be more popular across catalogues as
well. Therefore, we allow popular products to have more recommendations across
the two catalogues, proportional to their popularity.

In all, these popularity additions significantly boost performance in the
experiment. The results of this, along with the other modifications outlined in
this section, can be seen in the following section on performance.

\section*{Performance}
\subsection*{Baselines}
As in product-product content-similarity, it is difficult to capture similarity using
descriptions alone. It is not even clear that a mapping necessarily exists
between two seperately learned concept spaces. The concepts may represent
fundamentally different things, which cannot be well mapped to each other. We
can measure the effectiveness of the mapping by comparing it to a couple of
baselines. The upper limit = One Model, lower limit = random mapping.
Indeed, we find that this is a bottle neck of the operation.

Baseline popularity methods
[Number of products mapped to in different methods]

Random, tfidf, one model (explain), most popular

@@

\section*{Appendix}
